---
title: "P8105 HW3"
author: "Wenbo Fei"
date: "10/10/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(p8105.datasets)
knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

#Problem 1

```{r P1}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.  Each observation is an item specific record. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. For example, the first order record shows that in order1 made by user112108 at 10:00 on the forth day of a week after 9 days of his last purchase while this is his 4th order, a product "Bulgarian Yogurt" with ID 49302 is bought; this is an reodered item and it was the first one to be added in his cart; it belongs to dairy eggs department(16) on the yogurt aisle(120).

```{r P1Q1}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```

There are 134 aisles, and fresh vegetables are the most items ordered from.

Let's see the plot about number of items ordered in each aisle.

```{r P1Q2}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs( y = "number of items ordered")
```

Let's see a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.

```{r P1Q3}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```

Let's see a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r P1Q4}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```

#Problem 2
```{r P2_load_data, message=FALSE}
accel_data = 
  read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute_of_day",
    names_prefix = "activity_",
    values_to = "activity_count" ) %>%
    mutate( weekend = ifelse(day %in% c("Saturday","Sunday"), 1, 0),
            day = forcats::fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>%
  mutate_at(vars(week:day_id), as.factor) %>%
  mutate_at("minute_of_day", as.integer) %>%
  mutate_at("weekend", as.logical) 
  
head(accel_data)
```

Our final dataset includes `r nrow(accel_data)` observations with `r ncol(accel_data)` variables. Each observation has a "week" variable indicating which week, a unique"day_id" variable for each specific day, a "day" variable indicating the day of the week, a "minute_of_day" variable indicating which minute of the day is the record about, a "activity_court" recoding the activity counts for the minute, and a "weekend" indicating whether it's a weekend or a weekday.

aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r P2Q2}
accel_data %>%
  group_by(day_id) %>%
  summarize(total_activity = sum(activity_count)) %>%
  knitr::kable()
```
I can't see any apparent increasing or decreasing trend from this table. Looks like except for 3 days, all the other days have same order of magnitude of total activity count with fluctuation. We may need further analysis and visulization to find systematic trend.

#Problem 3

